Writing a reproducible paper with R Markdown and Pagedown^[Corresponding adress: mail@paulcbauer.eu, camille.landesvatter@mzes.uni-mannheim.de]

### 2021-06-22 paper.Rmd
Smaller edits.

Formatting text as PDF is probably one of the most widespread standards in the scientific community, especially when it comes to submitting papers and similar documents. The traditional way to well-formatted and good-looking PDFs is often through LaTeX or Word.

@@

While web pages are usually single-page scrollable documents, pagedown uses the JavaScript library `Paged.js` which allows documents to be paginated with elements like headers, footers and everything a readable scientific paper will need. Additionally, pagedown documents are based on R Markdown.

@@

We assume that you are using R on a day-to-day basis and you may have even started to work in R Markdown. If you don't know what R Markdown is there are many great resources (e.g. watch this [short video](https://vimeo.com/178485416)). Also, there is a template for a scientific paper written in R markdown by one of us two (Paul), which follows the same idea and structure just as this review ([github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)). The template compiles information, operations, tips and tricks you will very likely need to create a well-formatted scientific paper with R markdown.

@@

* ...install [R](https://www.r-project.org/) and [Rstudio](https://www.rstudio.com/) (most recent versions) [@R2017; @Rstudio2015].
* ...install the "pagedown"-package from github using the code below.
```{r eval=FALSE, include=TRUE}
remotes::install_github('rstudio/pagedown')
```
* ...also install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra].
```{r eval=FALSE, include=TRUE}
install.packages(c("rmarkdown", "knitr", "kableExtra",
                   "stargazer", "modelsummary", "knitr"))
```
* ...download the 4 input files we created --- `paper.rmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing). Ignore the other files.
* ...also download the 4 styling files we created: `wp_paged.html`, `wp.css`, `wp-fonts.css` and `wp-pages.css`.
* ...store all 8 files from above together in one folder (and use this folder as your working directory later on)
* ...learn R and read about the other underlying components namely [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Latex](https://en.wikipedia.org/wiki/LaTeX).
* ...pagedown comes with several Rmd-templates (presentations, poster, thesis, etc.) and via this review we provide another template for a working paper style. If however you want to modify single aspects or create your own template, you will need to at least gain some basic skills in [CSS](https://www.w3schools.com/css/) and [HTML](https://www.w3schools.com/html/).

>>>>

Formatting text as PDF is probably one of the most widespread standards in the scientific community, especially when it comes to submitting papers and similar documents. The traditional way to well-formatted and good-looking PDFs is often through LaTeX or Word. However, if you have spent hours and hours debugging latex code (or getting it to run) you may be on the lookout for something new.

@@

While web pages are usually single-page scrollable documents, pagedown uses the JavaScript library `Paged.js` which allows documents to be paginated with elements like headers, footers and everything a readable scientific paper will need. Additionally, pagedown documents are based on R Markdown. In our view, Pagedown and the underlying technology may replace Latex in the long run.

@@

We assume that you are using R on a day-to-day basis and you may have even started to work in R Markdown. If you don't know what R Markdown is there are many great resources that you should use (e.g. watch this [short video](https://vimeo.com/178485416)). An older template [see @Bauer2018-dl; https://osf.io/q395s] on which this newer template is based, may provide a quick entry point to writing a reproducible with R Markdown and Latex.

@@

* ...install [R](https://www.r-project.org/) and [Rstudio](https://www.rstudio.com/) (most recent versions) [@R2017; @Rstudio2015].
* ...install the "pagedown"-package from github using the code below [@Xie2021-bi; @Xie2021-ls].
```{r eval=FALSE, include=TRUE}
remotes::install_github('rstudio/pagedown')
```
* ...also install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra].
```{r eval=FALSE, include=TRUE}
install.packages(c("rmarkdown", "knitr", "kableExtra",
                   "stargazer", "modelsummary", "knitr"))
```
* ...download the 4 input files we created --- `paper.rmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing). Ignore the other files.
* ...also download the 4 styling files we created: `wp_paged.html`, `wp.css`, `wp-fonts.css` and `wp-pages.css`.
* ...store all 8 files from above together in one folder (and use this folder as your working directory later on)
* ...learn R and read about the other underlying components namely [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Latex](https://en.wikipedia.org/wiki/LaTeX).
* ...pagedown comes with several Rmd-templates (presentations, poster, thesis, etc.) and via this review we provide another template for a working paper style. If however you want to modify single aspects or create your own template, you will need to at least gain some basic skills in [CSS](https://www.w3schools.com/css/) and [HTML](https://www.w3schools.com/html/).
